The Driving Forces of API Evolution

### 2015-06-02 thesis/paper.tex
Few changes

\subsection{APIs} \label{api_design}
%what is an API
An API offers an interface through which developers can use to access functionality of a system. APIs can be seen as having three interacting layers; the interfaces, the implementation and additional functionality. Typically, the interfaces provide the definitions of functions and data structures while the implementation realises those interfaces. The additional functionality exists as a middle layer that provides things such as error handling, third party libraries and additional auxiliary features.
%we consider APIs to be systems because x y z \cite{761853,5733835}
Given that there is a combination of different interacting elements that does not only concern interfaces to implementations, we would argue that an API can be described as a system \cite{761853,5733835}.
%Introduce different layers of an API.
%Headers / Interfaces
%Implementation
%"Middle Layer"

\subsubsection{API Design}
API design is notoriously difficult, as a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating, bloch2008effective, stylos2006comparing}. Examples of such design decisions include how to structure an object's constructor parameters or if the API should display errors at compilation or at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user-defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design challenges, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API must be clear to the user \cite{shi2011empirical, stylos2006comparing}. The design decisions reached during development of an API will affect the overall usability of the API. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and its users \cite{clarke2004measuring}.

@@

Software evolution theory suggests that software programs can be grouped into three different types of systems, namely S-programs, P-programs and E-programs \cite{lehman1980programs}. S-programs are programs which are exactly derivable from a specification. An example of such a program is one which finds the lowest common denominator for two integers. A key characteristic of an S-program is that if its functionality is changed, the end product will be a new product and not an evolved version of the previous program, since the mapping between specification and program is exact. P-programs are programs which are inherently more complex, and ``solve a problem which can be precisely formulated, but whose solution must inevitably reflect an approximation of the real world''. Furthermore, P-programs are programs which are centred around the perception of users, analysts and/or programmers. This perception is often commonly a mathematical model of the problem. An example of such a program is a program which predicts weather, since it is derived from a set of hydrodynamic equations. E-programs are described as ``inherently even more change prone [than S-programs and P-programs]. They are programs that mechanize a human or societal activity''. These factors can be explained by the difference between P-programs and a E-programs which is that E-programs ``change the very nature of the problem to be solved'' by ``becoming a part of the world it models''. Further, a characteristic of E-programs is that, during development, the specification must be developed by involving predictions of the consequences of the deployment of that system will have, since once the program is deployed ``questions of correctness, appropriateness, and satisfaction arise which inevitably leads to additional pressure for change". Lehman's eight laws of software evolution are said to apply to such E-programs. The laws are presented in Table \ref{table:expl_laws}.

>>>>

\subsection{APIs} 
%what is an API
An API offers an interface through which developers can use to access functionality of a system. APIs can be seen as having three interacting layers; the interfaces, the implementation and additional functionality. Typically, the interfaces provide the definitions of functions and data structures while the implementation realises those interfaces. The additional functionality exists as a middle layer that provides utilities such as error handling, third party libraries and additional auxiliary features.
%we consider APIs to be systems because x y z \cite{761853,5733835}
Given that there is a combination of different interacting elements that does not only concern interfaces to implementations, we would argue that an API can be described as a system \cite{761853,5733835}.
%Introduce different layers of an API.
%Headers / Interfaces
%Implementation
%"Middle Layer"

\subsubsection{API Design} \label{api_design}
API design is notoriously difficult, as a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating, bloch2008effective, stylos2006comparing}. Examples of such design decisions include how to structure an object's constructor parameters or if the API should display errors at compilation or at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user-defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design challenges, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API must be clear to the user \cite{shi2011empirical, stylos2006comparing}. The design decisions reached during development of an API will affect the overall usability of the API. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and its users \cite{clarke2004measuring}.

@@

Software evolution theory suggests that software programs can be grouped into three different types of systems, namely S-programs, P-programs and E-programs \cite{lehman1980programs}. S-programs are programs which are exactly derivable from a specification. An example of such a program is one which finds the lowest common denominator for two integers. A key characteristic of an S-program is that if its functionality is changed, the end product will be a new product and not an evolved version of the previous program, since the mapping between specification and program is exact. P-programs are programs which are inherently more complex, and ``solve a problem which can be precisely formulated, but whose solution must inevitably reflect an approximation of the real world''. Furthermore, P-programs are programs which are centred around the perception of users, analysts or programmers. This perception is often commonly a mathematical model of the problem. An example of such a program is a program which predicts weather, since it is derived from a set of hydrodynamic equations. E-programs are described as ``inherently even more change prone [than S-programs and P-programs]. They are programs that mechanize a human or societal activity''. These factors can be explained by the difference between P-programs and a E-programs which is that E-programs ``change the very nature of the problem to be solved'' by ``becoming a part of the world it models''. Further, a characteristic of E-programs is that, during development, the specification must be developed by involving predictions of the consequences of the deployment of that system will have, since once the program is deployed ``questions of correctness, appropriateness, and satisfaction arise which inevitably leads to additional pressure for change". Lehman's eight laws of software evolution are said to apply to such E-programs. The laws are presented in Table \ref{table:expl_laws}.
